Best Practices for Alchemical Free Energy Calculations: v\versionnumber

### 2020-01-10 paper/manuscript.tex

\newcommand{\versionnumber}{0.1}  % you should update the minor version number in preprints and major version number of submissions.
\newcommand{\githubrepository}{\url{https://github.com/michellab/alchemical-best-practices}}  %this should be the main github repository for this article
\newcommand{\expect}[1]{\left\langle{#1}\right\rangle}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Best Practices for Alchemical Free Energy Calculations: v\versionnumber}
\author[2*]{Antonia S. J. S. Mey}
\author[7]{Bryce K. Allen}
\author[1*]{John D. Chodera}
\author[2,12]{Maximilian Kuhn}
\author[2*]{Julien Michel}
\author[3*]{David L. Mobley}
\author[1]{Levi N. Naden}
\author[4]{Samarjeet Prasad}
\author[5]{Julia E. Rice}
\author[1,8]{Andrea Rizzi}
\author[2]{Jenke Scheen}
\author[6*]{Michael R. Shirts}
\author[11]{Gary Tresadern}
\author[7]{Huafeng Xu}

\affil[1]{Computational and Systems Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York NY, USA}
\affil[2]{EaStCHEM School of Chemistry, David Brewster Road, Joseph Black Building, The King's Buildings, Edinburgh, EH9 3FJ, UK}
\affil[3]{Departments of Pharmaceutical Sciences and Chemistry, University of California, Irvine, USA}
\affil[4]{National Institutes of Health, Bethesda, MD, USA}
\affil[5]{IBM Alamden Research Center, Almaden, CA, USA}
\affil[6]{University of Colorado Boulder, Boulder, CO, USA}
\affil[7]{Silicon Therapeutics, Boston, MA, USA}
\affil[8]{Tri-Institutional Training Program in Computational Biology and Medicine, New York, NY, USA}
\affil[11]{Computational Chemistry, Janssen Research \& Development, Turnhoutseweg 30, Beerse B-2340,Belgium}
\affil[12]{Cresset, Cambridgeshire, UK}

\corr{john.chodera@choderalab.org}{JDC}
\corr{dmobley@mobleylab.org}{DLM}
\corr{antonia.mey@ed.ac.uk}{ASJSM}
\corr{mail@julienmichel.net}{JM}
\corr{michael.shirts@colorado.edu}{MRS}

\contrib[\authfn{1}]{These authors contributed equally to this work}
\contrib[\authfn{2}]{These authors also contributed equally to this work}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PUBLICATION INFORMATION
%%% Fill out these parameters when available
%%% These are used when the "pubversion" option is invoked
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pubDOI{10.XXXX/YYYYYYY}
\pubvolume{<volume>}
\pubyear{<year>}
\articlenum{<number>}
\datereceived{Month, Day, Year}
\dateaccepted{Month, Day, Year}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%
\begin{frontmatter}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Abstract
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Alchemical free energy calculations are a useful tool for predicting free energy differences associated with the transfer of small molecules from one environment to another.
The hallmark of these methods is the use of "bridging" potential energy functions representing \emph{alchemical} intermediate states that do not exist in as real chemical species. The data collected from these bridging alchemical thermodynamic states allows the computation of transfer free energies (or differences in transfer free energies) with orders of magnitude less simulation time than simulating the transfer process directly. 
%
While these methods are highly flexible, care must be taken in avoiding common pitfalls to ensure that computed free energy differences can be robust and reproducible for the chosen force field, and that appropriate corrections are included to permit direct comparison with experimental data.
%
In this paper, we review current best practices for several popular application domains of alchemical free energy calculations, including relative and absolute small molecule binding free energy calculations to biomolecular targets.
\todo[inline, color=green!20]{JDC: Should we migrate this document to \url{https://github.com/alchemistry}? ASJSM: Let's migrate with the submitted version, but leave it here for now?}
\end{abstract}
\end{frontmatter}

\todototoc
\listoftodos

%%%%%%%%%%%%%%%%%%%%
%  Introduction    %
%%%%%%%%%%%%%%%%%%%%

@@

\section{What are alchemical free energy methods?}
\label{sec:intro}
\todo[inline, color={green!20}]{@everyone: Any further input in this?}
Alchemical free energy (AFE) calculations compute free energy differences associated with transfer processes, such as the binding of a small molecule to a receptor or the transfer of a small molecule from an aqeuous to apolar phase~\cite{zwanzig1954hightemperature}. 
The calculations use non-physical intermediate states where atoms are changed from one type to another or can be removed or inserted by turning on or off the interactions with their environment.

Fig.~\ref{fig:fig_what_is_alchemy} illustrates common free energy changes that may be difficult to compute with conventional computational methods, but are more readily tractable with AFE methods. In these types of simulations so-called \textit{alchemical states} permit converged calculations for large changes, such as the entire removal of a ligand from a binding site, as seen in fig.~\ref{fig:fig_what_is_alchemy} (E). This type of simulation is known as an absolute AFE calculation. Other typical problems tacked with AFE calculations can be as simple as conformational changes of a molecule, that exhibit a high barrier between these (fig.~\ref{fig:fig_what_is_alchemy} (A)), computation of partition coefficients, such as $\log P$ or $\log D$ (fig.~\ref{fig:fig_what_is_alchemy} (B))~\cite{rustenburg2016measuring, bosisio2016blinded}, and membrane insertions (fig.~\ref{fig:fig_what_is_alchemy} (C))~\cite{}. Then there are common types of AFE that estimate free energies of binding: the ligand stays the same, but a protein residue is alchemically mutated and the effect on the binding affinity is probed (fig.~\ref{fig:fig_what_is_alchemy} (D))\cite{hauser2018predicting,aldeghi2018accurate}, changes in protein thermostability~\cite{seeliger2010protein,gapsys2016insights,gapsys2016accurate,aldeghi2019accurate}, absolute binding free energies (fig.~\ref{fig:fig_what_is_alchemy} (E))~\cite{}, and relative free energies of binding where the difference of binding free energy between to similar ligands is evaluated (fig.~\ref{fig:fig_what_is_alchemy} (F))~\cite{}.

In order to obtain a free energy from these AFE simulations, historically two common approaches were used for the analysis: free energy perturbation (FEP) using the Zwanzig relation~\cite{zwanzig1954hightemperature} or Bennett's acceptance ratio~\cite{bennett1976efficient} and thermodynamic integration (TI) for which the theory dates back decades but with the first computational applications emerging in the 1980's and 90's~\cite{kirkwood1935statistical, jorgensen1985monte, kollman1993free, wong1986dynamics, merz1989free}. Important studies continued during the 2000's often leading to improved implementation of the methods in common simulation software~\cite{vanderspoel2005gromacs, mermelstein2018fast, wang2015accurate, hedges2019biosimspace}. This foundational work, combined with the methodological, technological, and hardware improvements of the last 5-10 years have led to an explosion of interest and direct commercial application of the technologies~\cite{}.

@@

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig1_what_is_alchemy/Figure.pdf}   
    \caption{\textbf{Illustration of common types of free energies differences that can be calculated using alchemical free energy methods.} A: Change in free energy due to a conformational change of the molecule. B: Partition coefficient such as $\log P$ or $\log D$ can be computed using alchemical free energy methods. Here, as an example the partition coefficient between methanol and water is shown. C: Free energy difference associated with the insertion of a molecule into a membrane. D: Effect of mutations of protein or host residues on free energies of binding.  E: Absolute free energies of binding of a small molecule to a host (e.g. protein), F: Relative free energy of binding of one molecule with respect to another, here toluene and benzyl alcohol.
    \label{fig:fig_what_is_alchemy}
    }
\end{figure}

@@

We assume the reader possesses a basic familiarity with the principles of molecular mechanics, molecular dynamics simulations, statistical mechanics, and the biophysics of protein-ligand association.
For readers unfamiliar with these concepts that have never set up and run an equilibrium molecular dynamics simulation, we would recommend the following texts for getting started with these topics first. 
    \item Statistical mechanics and thermodynamics background~\cite{hill1956statistical, alavi2011statistical}
    \item Understanding theory of how molecular simulations work~\cite{frenkel2002understanding}
    \item Docking reviews~\cite{yuriev2013latest, meng2011molecular} and software overview~\cite{pagadala2017software} and generating valid molecular structures ~\cite{loeffler2015fesetup}
    \item Getting started with setting up and running equilibrium molecular dynamics simulations: Gromacs tutorial~\cite{lemkul2018From}, Amber tutorial~\cite{amber}, HTMD: high-throughput Molecular Dynamics for Molecular Discovery~\cite{doerr2016htmd}*, or with VMD and NAMD~\cite{ribeiro2016easy}*
References marked with an asterisk (*) indicate nonessential reading or tutorials that provide further depth and understanding.

@@

U - Potential energy

@@

\begin{tcolorbox}[title=Common terminology, colback=green!10!white]
    ligand - small molecular binding to a larger host molecule such as a protein.\\ 
    lead optimization - process of designing a drug candidate from an initial lead compound, considering its affinity and various pharmacokinetic parameters.\\
    thermodynamic cycle - sequence of thermodynamic processes (like changes in temperature, volume or pressure) in which the initial and final state are equal.
\end{tcolorbox}

%%%%%%%%%%%%%%%%%%%%
% Theory basics    %
%%%%%%%%%%%%%%%%%%%%

\section{Statistical mechanics can tell us why alchemical free energy calculations work}
\label{sec:theory}
\todo[inline, color=green!20]{ASJSM: This section needs to be expanded and give a good introduction to the theoretical background without going into too much detail.}
Why would you want to run an alchemical free energy calculation and why do they work? Suppose you want to compute the binding affinity, or free energy of binding, of a ligand $L$ to a protein $P$, given by:
\begin{equation}
P+L\leftrightharpoons PL
\end{equation}
The equilibrium rate of binding is given by the ratio of concentrations of product $[PL]$ and reactants $[P]$, $[L]$:
\begin{equation}
 K_b^{\circ} = c^{\circ}\frac{[PL]}{[L][P]}.
\end{equation}
In order to define a reference state, the standard state is usually defined as $1 \mathrm{mol}/\mathrm{L}$ and $c^\circ$ is the standard state concentration. Thus, the binding free energy is given by:
\begin{equation}
    \Delta G_{\mathrm{bind}} = -RT\ln K_b^{\circ}.
\end{equation}
One way to estimate the binding rate is by computing the probability of finding the molecular system in the bound state or the unbound state, making use of partition function ratios, i.e.

@@

It is expected that AFE methods can provide a valuable tool for drug discovery, capable of outperforming alternative approaches. (CITE https://pubs.acs.org/doi/abs/10.1021/acs.jmedchem.6b01881). It has been stressed that successful application in industry requires more knowledge of the "domain of applicability", in other words, a greater understanding of the success and failure~\cite{sherborne2016collaborating}, which in turn requires robust protocols for preparing, submitting and analysing AFE calculations. In this regard, the issues mentioned in the previous section such as understanding the suitability and timescales to capture the structure activity relationships (SAR), and performing up-front tests of performance are all relevant to drug discovery applications. Without venturing too far into details of system setup, which is beyond the scope of this article, there are some points to be stressed. 

@@

Methods in experimental design have been applied to the construction of the perturbation maps.  Yang et al.~\cite{yang2019optimal} optimized the perturbation map by selecting a fixed number of calculations from the pairwise perturbations so that the resulting set of calculations minimize the total variance.  Xu~\cite{xu2019diffnet} optimized the perturbation map by allocating different amounts of simulation time to different pairwise perturbations so as to minimize the total variance, given the total simulation time of all the perturbation calculations.  Both approaches lead to substantial reduction in the statistical error of the estimated free energies.  
%
\paragraph{Constraints and relative free energy calculations.}
One issue which requires particular care is the use of constraints.
Commonly, bonds involving hydrogen are constrained to a fixed length using algorithms such as SHAKE or LINCS, allowing the use of longer timesteps~\cite{}.
However, in single topology relative free energy calculations, the atoms involved might be mutated to other atom types -- for example, in a mutation of methane to methanol, one hydrogen might become an oxygen atom.
Typical molecular dynamics engines are not set up to recognize this change, or at least not to correctly include contributions to the free energy from changing constraints/constraint length, so results for a transformation would usually be erroneous.
At present the most general solution to this problem is simply to avoid the use of constraints (and thus use a smaller timestep if necessary, usually of around 1 fs) in any relative free energy calculation involving a transformation of a constrained bond.
%
\subsubsection{Absolute free energy calculations must handle the standard state and use restraints}
\label{sec:standardstate-restraints}
%
\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig5_thermo_cyc_abs/Figure.pdf}
    \caption{\textbf{Absolute thermodynamic cycle}}
    \label{fig:fig_absolute_thermodynamic_cycle}
\end{figure}

Absolute free energy calculations involve completely removing the interactions between the ligand or solute and its environment, taking it to a non-interacting state that may or may not retain intramolecular non-bonded interactions.
This non-interacting state can then be shifted between environments (from the protein to water, or from one solution to another) without changing its free energy, and then interactions can be restored.

Absolute free energies are typically reported with respect to a specific reference or standard state, which effectively determines the arbitrary point at which the free energy is 0.
The role of the standard state is particularly evident with binding free energies, in which having a reference state allows us to obtain a well-defined partial molar free energy for the reaction
\begin{equation*}
\ce{AB <=> A + B}
\end{equation*}
through the well-known expression derived from the law of mass action
\begin{equation} \label{eq:DGfromKAB}
\Delta G = -RT ~ \ln \left( c^{\circ} K_{AB} \right)  = -RT ~ \ln\left( \frac{c^{\circ} C_{AB}}{C_A C_B} \right) ,
\end{equation}
where $R$ is the gas, $T$ is the temperature, $C_X$ is the equilibrium concentration of the chemical species $X$ in the reaction solvent, and the reference state concentration $c^{\circ}$ converts the binding constant $K_{AB}$ into a dimensionless quantity expressed in reference concentration units.
It should be noted that ignoring the term $c^{\circ}$ is equivalent to assuming a reference concentration of 1~D$^{-1}$, where D are the units used to express $K_{AB}$, and would thus cause the value of $\Delta G$ to vary with the choice of the units.
Typically, it is convenient to define a standard state at a constant pressure of 1~atm and where each chemical species (i.e., A, B, and AB) in the reaction solvent has a concentration of $c^{\circ}$~=~1~M~=~1~molecule/1660~\r{A}$^3$ but do not interact with other molecules of A, B, or AB.

@@

For solvation free energy calculations, handling the standard state is typically straightforward, and treating it correctly simply means ensuring that the non-interacting solute is taken to the same (or equivalent) final reference state in both environments, e.g. that the transformation involves a 1M to 1M equivalent transfer free energy (where the non-interacting solute still occupies essentially the same volume as the solute in the interacting system).
So typically in such cases no special care is required to ensure the correct standard state, as long as the \emph{experimental} data being analyzed uses the same standard state and if it does not, a simple entropic correction is needed.

However, for binding the situation is much more complex and requires special care.
Experimental absolute binding free energies are reported relative to a specific reference state -- a 1 M standard state -- which must also be used in calculations.
In practice, this has implications for how the calculations are done, as the reference concentration must enter the thermodynamic cycle employed.

Typically, to deal with both practical sampling issues and the standard state issue, restraints are employed in absolute binding free energy calculations to keep the ligand in a well defined volume as its interactions with the system are removed~\cite{gilson1997statisticalthermodynamic}.
This solves two problems.
First, if the ligand were not kept in a well-defined region, as its interactions were removed it might wander the system, perhaps quite slowly, and only inadequately sample the non-interacting or weakly interacting state -- yet adequate sampling of these states might be required for convergence.
So for practical purposes, the use of restraints can dramatically improve sampling as interactions are weakened and removed.
Second, if the ligand is not kept in a well-defined region then it is hard to determine how to link a computed binding free energy to the correct 1 M standard state.
In contrast, with restraints, the free energy of releasing the restrained ligand to a 1 M standard state can be computed analytically or numerically by solving the relevant integral, allowing the standard state to enter the thermodynamic cycle.

\paragraph{Several choices of restraints are possible.}
\todo[inline, color={blue!20}]{@author: Check references}
In practice, a variety of types of restraints are common, from simple harmonic distance restraints between the ligand and the protein~\cite{mobley2006use}, to flat-bottom restraints which work similarly but only exert a force if the ligand leaves a specific region~\cite{}.

Alternatively, a set of restraints proposed by Boresch have also commonly been employed, where all six rigid-body degrees of freedom governing the orientation of the ligand relative to the receptor are restrained~\cite{leitgeb2005alchemical}.
Further restraints, such as on the overall ligand RMSD have also been used~\cite{woo2005calculation}.

In principle, all of these forms will yield correct binding free energies in the limit of adequate sampling (if their effects and connection to the standard state are correctly handled) but they have different strengths and weaknesses.
For example, with more involved restraints, sampling at intermediate lambda values will usually not need to be as extensive but more computational effort must go to computing the restraining free energy.
Additionally, such restraints would typically keep the ligand from exploring alternative binding modes, which may be undesirable with Hamiltonian lambda exchange or expanded ensemble techniques where allowing the ligand to exchange binding modes when it is non-interacting could provide sampling benefits~\cite{wang2013identifying}.
Concretely, flat bottom restraints might allow a ligand to explore multiple binding sites, harmonic restraints multiple binding modes within a site, and Boresch restraints a single binding mode within a single site.
See additional discussion of the possibility of multiple binding modes below~\ref{sec:multiple_binding_modes}.

>>>>

\newcommand{\versionnumber}{0.1}  % you should update the minor version number in preprints and major version number of submissions.
\newcommand{\githubrepository}{\url{https://github.com/michellab/alchemical-best-practices}}  %this should be the main github repository for this article
\newcommand{\expect}[1]{\left\langle{#1}\right\rangle}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE SETUP
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Best Practices for Alchemical Free Energy Calculations: v\versionnumber}
\author[1*]{Antonia S. J. S. Mey}
\author[7]{Bryce K. Allen}
\author[2*]{John D. Chodera}
\author[1,10]{Maximilian Kuhn}
\author[1*]{Julien Michel}
\author[3*]{David L. Mobley}
\author[1]{Levi N. Naden}
\author[4]{Samarjeet Prasad}
\author[5]{Julia E. Rice}
\author[2,8]{Andrea Rizzi}
\author[1]{Jenke Scheen}
\author[6*]{Michael R. Shirts}
\author[9]{Gary Tresadern}
\author[7]{Huafeng Xu}
\affil[1]{EaStCHEM School of Chemistry, David Brewster Road, Joseph Black Building, The King's Buildings, Edinburgh, EH9 3FJ, UK}
\affil[2]{Computational and Systems Biology Program, Sloan Kettering Institute, Memorial Sloan Kettering Cancer Center, New York NY, USA}
\affil[3]{Departments of Pharmaceutical Sciences and Chemistry, University of California, Irvine, USA}
\affil[4]{National Institutes of Health, Bethesda, MD, USA}
\affil[5]{IBM Alamden Research Center, Almaden, CA, USA}
\affil[6]{University of Colorado Boulder, Boulder, CO, USA}
\affil[7]{Silicon Therapeutics, Boston, MA, USA}
\affil[8]{Tri-Institutional Training Program in Computational Biology and Medicine, New York, NY, USA}
\affil[9]{Computational Chemistry, Janssen Research \& Development, Turnhoutseweg 30, Beerse B-2340,Belgium}
\affil[10]{Cresset, Cambridgeshire, UK}
\corr{john.chodera@choderalab.org}{JDC}
\corr{dmobley@mobleylab.org}{DLM}
\corr{antonia.mey@ed.ac.uk}{ASJSM}
\corr{mail@julienmichel.net}{JM}
\corr{michael.shirts@colorado.edu}{MRS}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% PUBLICATION INFORMATION
%%% Fill out these parameters when available
%%% These are used when the "pubversion" option is invoked
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pubDOI{10.XXXX/YYYYYYY}
\pubvolume{<volume>}
\pubyear{<year>}
\articlenum{<number>}
\datereceived{Month, Day, Year}
\dateaccepted{Month, Day, Year}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% ARTICLE START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\begin{frontmatter}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Abstract
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Alchemical free energy calculations are a useful tool for predicting free energy differences associated with the transfer of small molecules from one environment to another.
The hallmark of these methods is the use of "bridging" potential energy functions representing \emph{alchemical} intermediate states that do not exist in as real chemical species. The data collected from these bridging alchemical thermodynamic states allows the computation of transfer free energies (or differences in transfer free energies) with orders of magnitude less simulation time than simulating the transfer process directly. 
While these methods are highly flexible, care must be taken in avoiding common pitfalls to ensure that computed free energy differences can be robust and reproducible for the chosen force field, and that appropriate corrections are included to permit direct comparison with experimental data.
In this paper, we review current best practices for several popular application domains of alchemical free energy calculations, including relative and absolute small molecule binding free energy calculations to biomolecular targets.
\todo[inline, color=green!20]{JDC: Should we migrate this document to \url{https://github.com/alchemistry}? ASJSM: Let's migrate with the submitted version, but leave it here for now?}
\end{abstract}
\end{frontmatter}
\todototoc
\listoftodos
%  Introduction    %

@@

\section{What are alchemical free energy methods?}
\label{sec:intro}
\todo[inline, color={green!20}]{@everyone: Any further input in this?}
Alchemical free energy (AFE) calculations compute free energy differences associated with transfer processes, such as the binding of a small molecule to a receptor or the transfer of a small molecule from an aqueous to apolar phase~\cite{zwanzig1954hightemperature}. 
The calculations use non-physical intermediate states where atoms are changed from one type to another or can be removed or inserted by turning on or off the interactions with their environment. 
Fig.~\ref{fig:fig_what_is_alchemy} illustrates common free energy changes that may be difficult to compute with conventional computational methods, but are more readily tractable with AFE methods. In these types of simulations so-called \textit{alchemical states} permit converged calculations for large changes, such as the entire removal of a ligand from a binding site, as seen in fig.~\ref{fig:fig_what_is_alchemy} (E). This type of simulation is known as an absolute AFE calculation. Other typical problems tackled with AFE calculations can be as simple as conformational changes of a molecule that exhibit a high barrier between these (fig.~\ref{fig:fig_what_is_alchemy} (A)); computation of partition coefficients, such as $\log P$ or $\log D$ (fig.~\ref{fig:fig_what_is_alchemy} (B))~\cite{rustenburg2016measuring, bosisio2016blinded}; and membrane insertions (fig.~\ref{fig:fig_what_is_alchemy} (C))~\cite{}. Then there are common types of AFE that estimate free energies of binding: the ligand stays the same, but a protein residue is alchemically mutated and the effect on the binding affinity is probed (fig.~\ref{fig:fig_what_is_alchemy} (D))\cite{hauser2018predicting,aldeghi2018accurate}, changes in protein thermostability~\cite{seeliger2010protein,gapsys2016insights,gapsys2016accurate,aldeghi2019accurate}, absolute binding free energies (fig.~\ref{fig:fig_what_is_alchemy} (E))~\cite{}, and relative free energies of binding where the difference of binding free energy between to similar ligands is evaluated (fig.~\ref{fig:fig_what_is_alchemy} (F))~\cite{}.
In order to obtain a free energy from these AFE simulations, historically two common approaches were used for the analysis: free energy perturbation (FEP) using the Zwanzig relation~\cite{zwanzig1954hightemperature} or Bennett's acceptance ratio~\cite{bennett1976efficient} and thermodynamic integration (TI) for which the theory dates back decades but with the first computational applications emerging in the 1980's and 90's~\cite{kirkwood1935statistical, jorgensen1985monte, kollman1993free, wong1986dynamics, merz1989free}. Important studies continued during the 2000's often leading to improved implementation of the methods in common simulation software~\cite{vanderspoel2005gromacs, mermelstein2018fast, wang2015accurate, hedges2019biosimspace}. This foundational work, combined with the methodological, technological, and hardware improvements of the last 5-10 years have led to an explosion of interest and direct commercial application of the technologies~\cite{wang2015accurate, fratev2019improved, schindler2020largescale, cournia2017relative, sherborne2016collaborating}.

@@

\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig1_what_is_alchemy/Figure.pdf}   
    \caption{\textbf{Illustration of common types of free energies differences that can be calculated using alchemical free energy methods.} A: Change in free energy due to a conformational change of the molecule. B: Partition coefficient such as $\log P$ or $\log D$ can be computed using alchemical free energy methods. Here, as an example the partition coefficient between methanol and water is shown. C: Free energy difference associated with the insertion of a molecule into a membrane. D: Effect of mutations of protein or host residues on free energies of binding.  E: Absolute free energies of binding of a small molecule to a host (e.g. protein), F: Relative free energy of binding of one molecule with respect to another, here toluene and benzyl alcohol.
    \label{fig:fig_what_is_alchemy}
    }

@@

We assume the reader possesses a basic familiarity with the principles of molecular mechanics, molecular dynamics simulations, statistical mechanics, and the biophysics of protein-ligand association. If you feel unfamiliar with some of these concepts please good starting points can be found in these references~\cite{braun2019best, grossfield2018best, klimovich2015guidelines, shirts2012best}. Oftentimes, docking calculations are required to generated initial binding pose for a small molecule for an AFE calculation, where binding poses are not known, thus some basic familiarty on how to run docking calculations is also expected~\cite{grinter2014challenges}. 

@@

U - Potential energy\\

@@

\begin{tcolorbox}[title=Common terminology, colback=green!10!white]
    ligand - small molecular binding to a larger host molecule such as a protein.\\ 
    lead optimization - process of designing a drug candidate from an initial lead compound, considering its affinity and various pharmacokinetic parameters.\\
    thermodynamic cycle - sequence of thermodynamic processes (like changes in temperature, volume or pressure) in which the initial and final state are equal.\\
% Theory basics    %
\section{Statistical mechanics can tell us why alchemical free energy calculations work}
\label{sec:theory}
\todo[inline, color=green!20]{ASJSM: This section needs to be expanded and give a good introduction to the theoretical background without going into too much detail.}
Why would you want to run an alchemical free energy calculation and why do they work? Suppose you want to compute the binding affinity, or free energy of binding, of a ligand $L$ to a protein $P$, given by:
P+L\leftrightharpoons PL
The equilibrium rate of binding is given by the ratio of concentrations of product $[PL]$ and reactants $[P]$, $[L]$:
 K_b^{\circ} = c^{\circ}\frac{[PL]}{[L][P]}.
In order to define a reference state, the standard state is usually defined as $1 \mathrm{mol}/\mathrm{L}$ and $c^\circ$ is the standard state concentration. Thus, the binding free energy is given by:
    \Delta G_{\mathrm{bind}} = -RT\ln K_b^{\circ}.
One way to estimate the binding rate is by computing the probability of finding the molecular system in the bound state or the unbound state, making use of partition function ratios, i.e.

@@

It is expected that AFE methods can provide a valuable tool for drug discovery, capable of outperforming alternative approaches~\cite{kuhn2017prospective}. It has been stressed that successful application in industry requires more knowledge of the "domain of applicability", in other words, a greater understanding of the success and failure~\cite{sherborne2016collaborating}, which in turn requires robust protocols for preparing, submitting and analysing AFE calculations. In this regard, the issues mentioned in the previous section such as understanding the suitability and timescales to capture the structure activity relationships (SAR), and performing up-front tests of performance are all relevant to drug discovery applications. Without venturing too far into details of system setup, which is beyond the scope of this article, there are some points to be stressed. 

@@

Methods in experimental design have been applied to the construction of the perturbation maps.  Yang et al.~\cite{yang2019optimal} optimized the perturbation map by selecting a fixed number of calculations from the pairwise perturbations so that the resulting set of calculations minimize the total variance.  Xu~\cite{xu2019diffnet} optimized the perturbation map by allocating different amounts of simulation time to different pairwise perturbations so as to minimize the total variance, given the total simulation time of all the perturbation calculations.  Both approaches lead to substantial reduction in the statistical error of the estimated free energies.  
\paragraph{Constraints and relative free energy calculations.}
One issue which requires particular care is the use of constraints.
Commonly, bonds involving hydrogen are constrained to a fixed length using algorithms such as SHAKE or LINCS, allowing the use of longer timesteps~\cite{}.
However, in single topology relative free energy calculations, the atoms involved might be mutated to other atom types -- for example, in a mutation of methane to methanol, one hydrogen might become an oxygen atom.
Typical molecular dynamics engines are not set up to recognize this change, or at least not to correctly include contributions to the free energy from changing constraints/constraint length, so results for a transformation would usually be erroneous.
At present the most general solution to this problem is simply to avoid the use of constraints (and thus use a smaller timestep if necessary, usually of around 1 fs) in any relative free energy calculation involving a transformation of a constrained bond.
\subsubsection{Absolute free energy calculations must handle the standard state and use restraints}
\label{sec:standardstate-restraints}
\begin{figure}
    \includegraphics[width=0.95\linewidth]{figures/fig5_thermo_cyc_abs/Figure.pdf}
    \caption{\textbf{Absolute thermodynamic cycle}}
    \label{fig:fig_absolute_thermodynamic_cycle}
Absolute free energy calculations involve completely removing the interactions between the ligand or solute and its environment, taking it to a non-interacting state that may or may not retain intramolecular non-bonded interactions.
This non-interacting state can then be shifted between environments (from the protein to water, or from one solution to another) without changing its free energy, and then interactions can be restored.
Absolute free energies are typically reported with respect to a specific reference or standard state, which effectively determines the arbitrary point at which the free energy is 0.
The role of the standard state is particularly evident with binding free energies, in which having a reference state allows us to obtain a well-defined partial molar free energy for the reaction
\begin{equation*}
\ce{AB <=> A + B}
\end{equation*}
through the well-known expression derived from the law of mass action
\begin{equation} \label{eq:DGfromKAB}
\Delta G = -RT ~ \ln \left( c^{\circ} K_{AB} \right)  = -RT ~ \ln\left( \frac{c^{\circ} C_{AB}}{C_A C_B} \right) ,
where $R$ is the gas, $T$ is the temperature, $C_X$ is the equilibrium concentration of the chemical species $X$ in the reaction solvent, and the reference state concentration $c^{\circ}$ converts the binding constant $K_{AB}$ into a dimensionless quantity expressed in reference concentration units.
It should be noted that ignoring the term $c^{\circ}$ is equivalent to assuming a reference concentration of 1~D$^{-1}$, where D are the units used to express $K_{AB}$, and would thus cause the value of $\Delta G$ to vary with the choice of the units.
Typically, it is convenient to define a standard state at a constant pressure of 1~atm and where each chemical species (i.e., A, B, and AB) in the reaction solvent has a concentration of $c^{\circ}$~=~1~M~=~1~molecule/1660~\r{A}$^3$ but do not interact with other molecules of A, B, or AB.

@@

For solvation free energy calculations, handling the standard state is typically straightforward, and treating it correctly simply means ensuring that the non-interacting solute is taken to the same (or equivalent) final reference state in both environments, e.g. that the transformation involves a 1M to 1M equivalent transfer free energy (where the non-interacting solute still occupies essentially the same volume as the solute in the interacting system).
So typically in such cases no special care is required to ensure the correct standard state, as long as the \emph{experimental} data being analyzed uses the same standard state and if it does not, a simple entropic correction is needed.
However, for binding the situation is much more complex and requires special care.
Experimental absolute binding free energies are reported relative to a specific reference state -- a 1 M standard state -- which must also be used in calculations.
In practice, this has implications for how the calculations are done, as the reference concentration must enter the thermodynamic cycle employed.
Typically, to deal with both practical sampling issues and the standard state issue, restraints are employed in absolute binding free energy calculations to keep the ligand in a well defined volume as its interactions with the system are removed~\cite{gilson1997statisticalthermodynamic}.
This solves two problems.
First, if the ligand were not kept in a well-defined region, as its interactions were removed it might wander the system, perhaps quite slowly, and only inadequately sample the non-interacting or weakly interacting state -- yet adequate sampling of these states might be required for convergence.
So for practical purposes, the use of restraints can dramatically improve sampling as interactions are weakened and removed.
Second, if the ligand is not kept in a well-defined region then it is hard to determine how to link a computed binding free energy to the correct 1 M standard state.
In contrast, with restraints, the free energy of releasing the restrained ligand to a 1 M standard state can be computed analytically or numerically by solving the relevant integral, allowing the standard state to enter the thermodynamic cycle.
\paragraph{Several choices of restraints are possible.}
\todo[inline, color={blue!20}]{@author: Check references}
In practice, a variety of types of restraints are common, from simple harmonic distance restraints between the ligand and the protein~\cite{mobley2006use}, to flat-bottom restraints which work similarly but only exert a force if the ligand leaves a specific region~\cite{}.
Alternatively, a set of restraints proposed by Boresch have also commonly been employed, where all six rigid-body degrees of freedom governing the orientation of the ligand relative to the receptor are restrained~\cite{leitgeb2005alchemical}.
Further restraints, such as on the overall ligand RMSD have also been used~\cite{woo2005calculation}.
In principle, all of these forms will yield correct binding free energies in the limit of adequate sampling (if their effects and connection to the standard state are correctly handled) but they have different strengths and weaknesses.
For example, with more involved restraints, sampling at intermediate lambda values will usually not need to be as extensive but more computational effort must go to computing the restraining free energy.
Additionally, such restraints would typically keep the ligand from exploring alternative binding modes, which may be undesirable with Hamiltonian lambda exchange or expanded ensemble techniques where allowing the ligand to exchange binding modes when it is non-interacting could provide sampling benefits~\cite{wang2013identifying}.
Concretely, flat bottom restraints might allow a ligand to explore multiple binding sites, harmonic restraints multiple binding modes within a site, and Boresch restraints a single binding mode within a single site.
See additional discussion of the possibility of multiple binding modes below~\ref{sec:multiple_binding_modes}.
Many choices of restraints involve selecting reference atoms.
